### 2023-05-16 03_real-dends.qmd

---
    title: "Discovering Focused Microgenre Communities: Dendrograms"
    author: "Omar Lizardo"
    date: "`r Sys.Date()`"
    output:
        html_document: default
---

# Setup
```{r setup}
    knitr::opts_chunk$set(include=FALSE, echo = TRUE, warning = FALSE, message = FALSE)
    require("dendextend")
    require("here")
    require("RColorBrewer")
```

```{r Loading link distance matrix and performing link clustering}
    load(file= here("dat", "lclust-res.RData"))
```

# Dendrogram Plots from Link Clustered Data
```{r Plotting macrogenre dendrograms (1)}
    k <- length(dend.macro$lower)
    png(here("Plots", "Dend", "all-branches-macro.png"), height = 900, width = 1600)
    d <- lclust.dend
    d %>% 
      set("labels", c(rep("", nleaves(d)))) %>% 
      set("branches_k_color", value = brewer.pal(k, "Dark2"), k = k) %>% 
      set("branches_lwd", value = 3) %>%
      plot(axes = FALSE) %>% 
      abline(h = 8, col="red", lty = 2, lwd = 3) 
```

```{r Plotting macrogenre dendrograms (2)}
    k <- length(dend.macro$lower)
    png(here("Plots", "Dend", "all-branches-macro-labels.png"), height = 900, width = 1600)
    d <- dend.macro$upper
    d %>% 
      set("labels", dend.macro$lower %>% names) %>% 
      set("branches_k_color", value = brewer.pal(k, "Dark2")) %>% 
      set("branches_lwd", value = 4) %>% 
      set("labels_color", value = brewer.pal(k, "Dark2")) %>% 
      set("labels_cex", value = 2.5) %>% 
      plot(axes = FALSE) 
    abline(h = 8, col="red", lty = 2, lwd = 3)
```

```{r Plotting microgenre dendrograms (1)}
    k <- length(dend.micro$lower)
    png(here("Plots", "Dend", "all-branches-micro.png"), height = 900, width = 1600)
    k <- length(dend.micro$lower)
    d <- lclust.dend
    d %>% 
      set("labels", c(rep("", nleaves(d)))) %>% 
      set("branches_k_color", value = brewer.pal(k, "Dark2"), k = k) %>% 
      set("branches_lwd", value = 3) %>%
      plot(axes = FALSE) %>% 
      abline(h = 3, col="red", lty = 2, lwd = 3) 
```

```{r Plotting microgenre dendrograms (2)}
    k <- length(dend.micro$lower)
    png(here("Plots", "Dend", "all-branches-micro-labels.png"), height = 900, width = 1600)
    d <- dend.micro$upper
    d %>% 
      set("labels", micro.g.names) %>% 
      set("branches_k_color", value = brewer.pal(k, "Dark2"), k = k) %>% 
      set("branches_lwd", value = 3) %>% 
      set("labels_color", value = brewer.pal(k, "Dark2")) %>% 
      set("labels_cex", value = 0.9) %>% 
      plot(axes = FALSE) 
    par(cex= 2.5)
    title(xlab="", ylab="", main="") %>% 
      abline(h = 3, col="red", lty = 2, lwd = 3)
```

```{r Plotting genre-specific dendrograms (1)}
    png(here("Plots", "Dend", "classic-rock-branches.png"), height = 900, width = 1600)
    k <- n.micro["Classic_Rock"]
    d <- dend.macro$lower$Classic_Rock
    d %>% 
      set("labels", c(rep("", nleaves(d)))) %>% 
      set("branches_k_color", value = brewer.pal(k, "Dark2"), k = k) %>% 
      set("branches_lwd", value = 3) %>% 
      plot(axes = FALSE) 
    par(cex = 5)
    title(xlab="", ylab="", main="Classic Rock") %>% 
      abline(h = 3, col="red", lty = 2, lwd = 3)
```

```{r Plotting genre-specific dendrograms (2)}
    png(here("Plots", "Dend", "blues-branches.png"), height = 900, width = 1600)
    k <- n.micro["Blues_RandB"]
    d <- dend.macro$lower$Blues_RandB
    d %>% 
      set("labels", c(rep("", nleaves(d)))) %>% 
      set("branches_k_color", value = brewer.pal(k, "Dark2"), k = k) %>% 
      set("branches_lwd", value = 3) %>%        
      plot(axes = FALSE) 
    par(cex = 5)
    title(xlab="", ylab="", main="Blues/R&B") %>% 
      abline(h = 3, col="red", lty = 2, lwd = 3)
```

```{r Plotting genre-specific dendrograms (3)}
    png(here("Plots", "Dend", "opera-branches.png"), height = 900, width = 1600)
    k <- n.micro["Opera"]
    d <- dend.macro$lower$Opera
    d %>% 
      set("labels", c(rep("", nleaves(d)))) %>% 
      set("branches_k_color", value = brewer.pal(k, "Dark2"), k = k) %>% 
      set("branches_lwd", value = 3) %>%
      plot(axes = FALSE) 
    par(cex = 5)
    title(xlab="", ylab="", main="Opera") %>% 
      abline(h = 3, col="red", lty = 2, lwd = 3)
```