### 2023-07-31 01_link-distance-matrix.qmd
Updated linkdist function

---
    title: "Discovering Focused Microgenre Communities: Link Clustering"
    author: "Omar Lizardo"
    date: "`r Sys.Date()`"
    output:
        html_document: default
---

# Setup
```{r setup}
    knitr::opts_chunk$set(include=FALSE, echo = TRUE, warning = FALSE, message = FALSE)
    library(conflicted)
    library(dendextend)
    library(fastcluster)
    library(here)   
    library(haven)
    library(tidyverse)
```

# Data Wrangling
```{r Importing and creating two mode data frame of people by genres}
    taste.dat <- read_dta("C:/Users/Omar Lizardo/Google Drive/MISC DATA SOURCES/SSI-2012/SSI2012.dta")
    taste.dat <- taste.dat %>% 
      dplyr::select("id", ends_with(c("lis", "like")), -starts_with("none")) %>% 
      dplyr::select(c(1:41)) %>% 
      na.omit() %>% 
      mutate(Classical = classicallike * classicallis,
             Opera = operalike * operalis,
             Jazz = jazzlike * jazzlis,
             Broadway = bwaystlike * bwaystlis,
             Easy = moodezlike * moodezlis, 
             Bigband = bbandlike * bbandlis,
             Classic_Rock = croldlike * croldlis,
             Country = countrylike * countrylis,
             Bluegrass = blueglike * blueglis,
             Folk = folklike * folklis,
             Gospel = hymgoslike * hymgoslis,
             Latin = latlpsallike * latspsallis,
             Rap_Hip_Hop = raphiphoplike * raphiphoplis,
             Blues_RandB = blurblike * blurblis,
             Reggae = reggaelike * reggaelis,
             Pop = toppoplike * toppoplis,
             Contemp_Rock = controcklike * controcklis,
             Indie_Alt = indaltlike * indaltlis,
             Dance_Club = danclublike * danclublis,
             Heavy_Metal = hvymtllike * hvymtllis
             ) %>%  #people are linked to genres that the both like and listen to
      dplyr::select(id, Classical:Heavy_Metal)
      taste.mat <- as.matrix(taste.dat[, 2:ncol(taste.dat)])
```

# Link Clustering
```{r Creating link by link distance matrix and saving results}
    source(here("Functions", "link.dist.R"))
    ldist.res <- link.dist(x = taste.mat, id = taste.dat$id)
    b <- ls()[! ls() %in% "ldist.res"]
    rm(list = b)
    ls()
    save.image(file= here("dat", "lclust-res.RData"))
```