### 2023-08-03 qmd/02_toy-ex.qmd

---
title: "Discovering Focused Microgenre Communities: Toy example"
author: "Omar Lizardo"
date: "`r Sys.Date()`"
---

# Setup
```{r setup}
    knitr::opts_chunk$set(include=FALSE, echo = TRUE, warning = FALSE, message = FALSE)
    library("conflicted")
    library("factoextra")
    library("here")
    library("ggplot2")
    library("sjPlot")
    source(here("Functions", "link.dist.R"))
```

# Link Clustering (Toy Example)
```{r Link Clustering toy example}
    x <- c(0, 1, 1, 0, 
           0, 0, 0, 1, 
           1, 1, 1, 1, 
           1, 0, 0, 1, 
           0, 0, 1, 1)
    x <- matrix(x, nrow = 5, ncol = 4,  byrow = TRUE) #toy data
    ldist.res <- link.dist(x = x, id = c(1:5)) #link distance matrix
    hc.res <- hclust(ldist.res$dist, method = "ward.D2") #link clustering
```

```{r Plotting toy example dendrogram}
    hc.res$labels <- c("P1-G2", "P1-G3", "P2-G4", "P3-G1", "P3-G2", "P3-G3", "P3-G4", "P4-G1", "P4-G4", "P5-G3", "P5-G4")
    p <- fviz_dend(hc.res, k = 4, k_colors = "uchicago", show_labels = TRUE, cex = 1.5, lwd = 2) 
    p <- p + geom_hline(yintercept = 0.75, 
                        linetype = "dashed", 
                        color = "red", 
                        size = 1.2)
    p <- p + labs(y = "", title = "") 
    p <- p + theme(axis.line.y = element_blank(), axis.text.y = element_blank())
    p
    save_plot(here("Plots", "Toy", "link-clust-toy5.png"), 
              width = 40, height = 25)
```