### 2023-08-03 qmd/04_microgenre-analysis.qmd

---
  title: "Discovering Focused Microgenre Communities: Microgenre Analysis"
  author: "Omar Lizardo"
  date: "`r Sys.Date()`"
  output:
    html_document: default
---

# Setup
```{r setup}
    knitr::opts_chunk$set(include=FALSE, echo = TRUE, warning = FALSE, message = FALSE)
    library(conflicted)
    library(dendextend)
    library(ggplot2)
    library(haven)
    library(here)
    library(sjPlot)
    library(tidyverse)
    library(webshot)
    load(file= here("dat", "lclust-res.RData"))
```

```{r Importing and creating two mode data frame of people by genres}
    taste.dat <- read_dta("C:/Users/Omar Lizardo/Google Drive/MISC DATA SOURCES/SSI-2012/SSI2012.dta")
    taste.dat <- taste.dat %>% 
      dplyr::select("id", ends_with(c("lis", "like")), -starts_with("none")) %>% 
      dplyr::select(c(1:41)) %>% 
      na.omit() %>% 
      mutate(Classical = classicallike * classicallis,
             Opera = operalike * operalis,
             Jazz = jazzlike * jazzlis,
             Broadway = bwaystlike * bwaystlis,
             Easy = moodezlike * moodezlis, 
             Bigband = bbandlike * bbandlis,
             Classic_Rock = croldlike * croldlis,
             Country = countrylike * countrylis,
             Bluegrass = blueglike * blueglis,
             Folk = folklike * folklis,
             Gospel = hymgoslike * hymgoslis,
             Latin = latlpsallike * latspsallis,
             Rap_Hip_Hop = raphiphoplike * raphiphoplis,
             Blues_RandB = blurblike * blurblis,
             Reggae = reggaelike * reggaelis,
             Pop = toppoplike * toppoplis,
             Contemp_Rock = controcklike * controcklis,
             Indie_Alt = indaltlike * indaltlis,
             Dance_Club = danclublike * danclublis,
             Heavy_Metal = hvymtllike * hvymtllis
             ) %>%  #people are linked to genres that the both like and listen to
      dplyr::select(id, Classical:Heavy_Metal)
```

```{r Performing Link Clustering}
    lclust.res <- hclust(ldist.res$dist, method = "ward.D2") #clustering of link by link distance matrixco
    lclust.dend <- as.dendrogram(lclust.res) #transforming link clustering into dendrogram
```

```{r Loading Custom Functions}
    source(here("Functions", "micro.genre.dat.R"))
    source(here("Functions", "demog.dat.R"))
    source(here("Functions", "genre.reg.mods.R"))
    source(here("Functions", "pred.probs.R"))
    source(here("Functions", "plot.preds.R"))
```

```{r Microgenre Regression Moldes}
    spec <- c("poly(educ, 2)", "parcoll", "poly(age.n, 2)", 
              "blk", "his", "asn", "gender")
    labs <- c("Intercept", "Education", "Parents College",
                            "Age", "Age-Squared", 
                            "Race ID: Black", "Race ID: Hispanic",
                            "Race ID: Asian", "Gender ID: Woman")

    res.latin <- genre.reg.mods("Latin", k = 4, xvars = spec)
    res.metal <- genre.reg.mods("Heavy_Metal", k = 4, xvars = spec)

    tab_model(res.latin$mods, 
              digits = 3, show.ci = FALSE, 
              auto.label = FALSE, transform = NULL,
              pred.labels = labs,
              file = here("Tabs", "reg-tab-latin.html"))
    webshot(here("Tabs", "reg-tab-latin.html"), here("Tabs", "reg-tab-latin.png"))

    tab_model(res.metal$mods, 
              digits = 3, show.ci = FALSE, 
              auto.label = FALSE, transform = NULL,
              pred.labels = labs,
              file = here("Tabs", "reg-tab-metal.html"))
    webshot(here("Tabs", "reg-tab-metal.html"), here("Tabs", "reg-tab-metal.png"))
```

```{r Microgenre Regression Plots (Age)}
    age.preds <- pred.probs(x = "age.n", res.latin$mods)
    p <- plot.preds(d = age.preds, 
                    xb = c(3, 7, 11),
                    xlabs = c("Young", "Mid. Aged", "Old"),
                    yl = .1, 
                    xm = 2, 
                    xl = 12)
    save_plot(filename = here("Plots", "Demog", "micro-by-age-latin.png"), 
              fig = p,
              height = 25, width = 20)

    age.preds <- pred.probs(x = "age.n", res.metal$mods)
    p <- plot.preds(d = age.preds, 
                    xb = c(3, 7, 11),
                    xlabs = c("Young", "Mid. Aged", "Old"),
                    yl = .1, 
                    xm = 2, 
                    xl = 12)
    save_plot(filename = here("Plots", "Demog", "micro-by-age-metal.png"), 
              fig = p,
              height = 25, width = 20)
```
`