### 2023-08-05 old_Rmds/Old/other-link-clustering-analyses.Rmd

```{r calculating micro-genre distance matrix from micro-genre overlap matrix}
    micro.overlaps <- t(as.matrix(lclust.dat[, 2:ncol(lclust.dat)])) %*% as.matrix(lclust.dat[, 2:ncol(lclust.dat)])
    micro.dist <- micro.overlaps 
    diag(micro.dist) <- 0
    micro.dist <- as.dist(1 - micro.dist/max(micro.dist))
    micro.clust.res <- hclust(micro.dist, method = "ward.D2") #link clustering of link by link distance matrix
    nc <- NbClust(diss = micro.dist, method = "ward.D2", 
                      index = "silhouette", distance = NULL)
```

```{r prepping data for mca plot of micro-genre communities}
    mca.dat <- lclust.dat[, 2:ncol(lclust.dat)]
    factor_commvars <- function(x) {
        factor(x, labels = c("No", "Yes"))
      }
    mca.dat <- mca.dat %>% 
        mutate(across(1:ncol(mca.dat), factor_commvars))
    mca.res <- MCA(mca.dat, ncp = 10, graph = FALSE,  method = "indicator") #peforming MCA of micro-genre communities
    mca.yes.mods <- mca.res$var$coord[seq(2, nrow(mca.res$var$coord), by = 2), ]
    mca.clust.dat <- as_tibble(mca.yes.mods)
    mca.hk.res <-hkmeans(mca.clust.dat[,1:4], k = 17, hc.method = "ward.D2") #clustering yes modalities on principal components
    mca.clust.dat <- bind_cols(
      tibble(id = gsub("_Yes", "", rownames(mca.yes.mods))),
      mca.clust.dat, 
      tibble(cluster = factor(mca.hk.res$cluster))
      )
```

```{r MCA biplot of micro-genre communities}
    p <- ggplot(data = mca.clust.dat, 
                mapping = aes(x = `Dim 1`, y = `Dim 2`, color = cluster))
    p <- p + theme_minimal()
    p <- p + geom_hline(yintercept = 0, colour = "gray75") 
    p <- p + geom_vline(xintercept = 0, colour = "gray75")
    p <- p + geom_text(aes(label = id), check_overlap = TRUE, size = 4)
    p <- p + stat_ellipse(linetype = 0, geom = "polygon", aes(fill = cluster), alpha = 0.02)
    p <- p + labs(x = "Dimension 1", y = "Dimension 2")
    p <- p + theme_sjplot()
    p.12 <- p + theme(
                    panel.grid.major.y = element_line(color = "grey90"),
                    panel.grid.major.x = element_line(color = "grey90"),
                    axis.title = element_text(size = 18), 
                    axis.text = element_text(size = 18),
                    axis.line.x = element_blank(),
                    axis.line.y = element_blank(),
                    legend.position = "none"
                    )
    p <- ggplot(data = mca.clust.dat, 
                mapping = aes(x = `Dim 3`, y = `Dim 4`, color = cluster))
    p <- p + theme_minimal()
    p <- p + geom_hline(yintercept = 0, colour = "gray75") 
    p <- p + geom_vline(xintercept = 0, colour = "gray75")
    p <- p + geom_text(aes(label = id), check_overlap = TRUE, size = 4)
    p <- p + stat_ellipse(linetype = 0, geom = "polygon", aes(fill = cluster), alpha = 0.02)
    p <- p + labs(x = "Dimension 3", y = "Dimension 4")
    p <- p + theme_sjplot()
    p.34 <- p + theme(
                    panel.grid.major.y = element_line(color = "grey90"),
                    panel.grid.major.x = element_line(color = "grey90"),
                    axis.title = element_text(size = 18), 
                    axis.text = element_text(size = 18),
                    axis.line.x = element_blank(),
                    axis.line.y = element_blank(),
                    legend.position = "none"
                    )
    p <- fviz_nbclust(mca.yes.mods, 
                      kmeans, method = "silhouette", 
                      k.max = 20, linecolor = "red")     
    p <- p + labs(y = "", x = "") 
    p.in <- p + theme(plot.background = 
                       element_rect(colour = "white", size = 0.25),
                   axis.text.x = element_text(size = 10))
    p <-  p.12 / p.34 #combining plots using patchwork
    p <- p + inset_element(p.in, 0.01, 0, 0.35, 0.35) #inset plot using patchwork
    save_plot(here("Plots", "Link Clust", "micro-genre-mca.png"), 
              width = 35, height = 45)

```

```{r Network plot of two mode micro-genre overlaps}
    linkcomm.net <- graph_from_adjacency_matrix(micro.overlaps, 
                                            mode = "undirected",
                                            diag = FALSE, weighted = TRUE)
    linkcomm.bb <- disparity(linkcomm.net, alpha = 0.05, narrative = TRUE)
    iso <- which(degree(linkcomm.bb)== 0)
    linkcomm.bb <- delete.vertices(linkcomm.bb, iso) 
    c <- cluster_louvain(linkcomm.bb)
    V(linkcomm.bb)$c <- c$membership   
    p <- ggraph(linkcomm.bb, layout = 'auto')
    p <- p + geom_node_text(aes(label = name, 
                                color = as.factor(c)), size = 6, check_overlap = TRUE)
    p <- p + geom_edge_link(color = "grey70", width = 0.5) 
    p <- p + theme_graph() + theme(legend.position = "none")
    p
    save_plot(here("Plots", "Link Clust", "micro-genre-net.png"), 
              width = 55, height = 45)
```

```{r Dendrogram representation of nested micro-genre communities}
    p <- fviz_dend(micro.clust.res, cex = 1.05, k = nc$Best.nc[1], 
                       type = "rectangle", repel = TRUE, k_colors = "uchicago",
                       color_labels_by_k = TRUE, rect = FALSE, lwd = 1.2) 
    p <- p + theme_minimal() 
    p <- p + labs(x = "", y = "")
    p.rect <- p + theme(
                axis.line=element_blank(),
                axis.text.x=element_blank(),
                axis.text.y=element_blank(),
                axis.ticks=element_blank(),
                axis.title.x=element_blank(),
                axis.title.y=element_blank()
                )
    p.rect
    save_plot(here("Plots", "Link Clust", "micro-genre-dend.png"), 
              width = 65, height = 45)
```

```{r Phylogenic representation of nested micro-genre communities}
    p <- fviz_dend(micro.clust.res, cex = 1.5, k = nc$Best.nc[1], 
                       type = "phylogenic", repel = TRUE, k_colors = "uchicago",
                       color_labels_by_k = TRUE) 
    p <- p + theme_minimal() 
    p <- p + labs(x = "", y = "")
    p.phylo <- p + theme(
                axis.line=element_blank(),
                axis.text.x=element_blank(),
                axis.text.y=element_blank(),
                axis.ticks=element_blank(),
                axis.title.x=element_blank(),
                axis.title.y=element_blank()
                )
    p.phylo
    save_plot(here("Plots", "Link Clust", "micro-genre-phylo.png"), 
              width = 55, height = 45)
```