### 2023-08-05 old_Rmds/Old/trad-analysis-latest.Rmd

---
title: "Discovering Focused Microgenre Communities"
author: "Omar Lizardo"
date: "`r Sys.Date()`"
output:
html_document: default
---

# Setup
```{r setup}
    knitr::opts_chunk$set(include=FALSE, echo = TRUE, warning = FALSE, message = FALSE)
    require("depmixS4")
    require("FactoMineR")
    require("factoextra")
    require("ggraph")
    require("ggplot2")
    require("haven")
    require("here")   
    require("patchwork")
    require("qgraph")
    require("sjPlot")
    require("tidyverse")
```

# Data Wrangling
```{r Importing and creating two mode data frame of people by genres}
    taste.dat <- read_dta("C:/Users/Omar Lizardo/Google Drive/MISC DATA SOURCES/SSI-2012/SSI2012.dta")
    taste.dat <- taste.dat %>% 
      dplyr::select("id", ends_with(c("lis", "like")), -starts_with("none")) %>% 
      dplyr::select(c(1:41)) %>% 
      na.omit() %>% 
      mutate(Classical = classicallike * classicallis,
             Opera = operalike * operalis,
             Jazz = jazzlike * jazzlis,
             Broadway = bwaystlike * bwaystlis,
             Easy = moodezlike * moodezlis, 
             Bigband = bbandlike * bbandlis,
             Classic_Rock = croldlike * croldlis,
             Country = countrylike * countrylis,
             Bluegrass = blueglike * blueglis,
             Folk = folklike * folklis,
             Gospel = hymgoslike * hymgoslis,
             Latin = latlpsallike * latspsallis,
             Rap_Hip_Hop = raphiphoplike * raphiphoplis,
             Blues_RandB = blurblike * blurblis,
             Reggae = reggaelike * reggaelis,
             Pop = toppoplike * toppoplis,
             Contemp_Rock = controcklike * controcklis,
             Indie_Alt = indaltlike * indaltlis,
             Dance_Club = danclublike * danclublis,
             Heavy_Metal = hvymtllike * hvymtllis
             ) %>%  #people are linked to genres that the both like and listen to
    dplyr::select(id, Classical:Heavy_Metal)
    g.names <- names(taste.dat[, 2:ncol(taste.dat)])

```

# Traditional Analytic Techniques
```{r Factor Analysis}
    f1 <- c(12, 13, 14, 15, 19)
    f2 <- c(1, 2, 3, 4, 5, 6)
    f3 <- c(7, 16, 17, 18, 20)
    f4 <- c(8, 9, 10, 11)
    gen.groups <- list(f1, f2, f3, f4)
    efa.res <- factanal(taste.dat[, g.names], 
                        factors = 4, rotation = "promax",
                        scores = "regression")
    png(here("Plots", "Regular", "genre-efa-regular.png"), 
        height = 3900, width = 6500)
    qgraph(efa.res$loadings, groups = gen.groups, borders = FALSE, 
          edge.labels = TRUE, minimum = 0.05, 
          edge.label.cex = 0.8, edge.width = 0.8)
```

```{r Latent Class Analysis}
    lca.dat <- taste.dat[, g.names]
    set.seed(25)
    mod1 <- mix(list(Classical~1, Opera~1, Jazz~1, Broadway~1, Easy~1, Bigband~1, Classic_Rock~1, Country~1, Bluegrass~1, Folk~1, Gospel~1, Latin~1, Rap_Hip_Hop~1, Blues_RandB~1, Reggae~1, Pop~1, Contemp_Rock~1, Indie_Alt~1, Dance_Club~1, Heavy_Metal~1), 
                data=lca.dat, # the dataset to use
                nstates = 6, # the number of latent classes
                family=list(multinomial("identity"), multinomial("identity"), 
                             multinomial("identity"), multinomial("identity"),
                             multinomial("identity"), multinomial("identity"), 
                             multinomial("identity"), multinomial("identity"),
                             multinomial("identity"), multinomial("identity"), 
                             multinomial("identity"), multinomial("identity"),
                             multinomial("identity"), multinomial("identity"), 
                             multinomial("identity"), multinomial("identity"),
                             multinomial("identity"), multinomial("identity"), 
                             multinomial("identity"), multinomial("identity")),
               respstart=runif(240))
    fmod1 <- fit(mod1, verbose=FALSE)
    lca.res <- as_tibble(posterior(fmod1)) #collecting latent class probabilities
```

```{r Data table for plotting latent classes}
    lc.plot.data <- bind_cols(lca.dat, lca.res) %>% 
        gather(key="measure", value="value", Classical:Heavy_Metal) %>% 
        mutate(state = factor(state, labels = c("Class 1", "Class 2", "Class 3", "Class 4", "Class 5", "Class 6"))) %>% 
        mutate(value = factor(value, labels = c("no", "yes")))
```

```{r Plotting latent classes}
    p <- ggplot(data = lc.plot.data, mapping = aes(x = measure, fill = value)) 
    p <- p + theme_sjplot()
    p <- p + geom_bar(position="fill")  
    p <- p + scale_fill_brewer(palette = "Paired")
    p <- p + facet_wrap(~ state, nrow = 2) + coord_flip() 
    p <- p + theme(
                axis.text = element_text(size = 12),
                legend.position = "bottom",
                strip.text = element_text(size = 18)
                )

    p.lca <- p + labs(x = "", y = "") + labs(fill="")
    p.lca
    save_plot(here("Plots", "Regular", "genre-lca-regular.png"), 
              width = 40, height = 25)
```

```{r Prepping data for MCA, performing MCA, and generating label size and individual alpha vectors}
    yesno.labs <- function(x) {
        factor(x, labels = c("No", "Yes"))
        }
    mca.dat <- taste.dat[, g.names] %>% 
        mutate(across(1:ncol(taste.dat)-1, yesno.labs))
    mca.res <- MCA(mca.dat, ncp = 10, graph = FALSE,  method = "indicator") #performing MCA
    mca.dim.vars <- as_tibble(mca.res$var$coord) #modality coordinates
    mca.dim.ind <- as_tibble(mca.res$ind$coord) #individual coordinates
    var.names <- rownames(mca.res$var$coord) #vector of label names
    text.size1 <- log(apply(mca.res$var$contrib[,c(1:2)], 1, sum))*3 #vector of sizes based on contributions
    text.size2 <- log(apply(mca.res$var$contrib[,c(3:4)], 1, sum))*3  #vector of sizes based on contributions
    ind.alpha1 <- apply(mca.res$ind$contrib[,c(1:2)], 1, mean)/ max(apply(mca.res$ind$contrib[,c(1:2)], 1, mean))  #vector of alphas based on contributions
    ind.alpha2<- apply(mca.res$ind$contrib[,c(3:4)], 1, mean)/ max(apply(mca.res$ind$contrib[,c(3:4)], 1, mean)) #vector of alphas based on contributions
```

```{r MCA biplots}
    p <- ggplot(data = mca.dim.vars, mapping = aes(x = `Dim 1`, y = `Dim 2`))
    p <- p + geom_text(aes(label = var.names), size = text.size1, color = "blue", check_overlap = TRUE) #cloud of modalities
    p <- p + geom_hline(yintercept = 0, colour = "gray45") 
    p <- p + geom_vline(xintercept = 0, colour = "gray45") 
    p <- p + labs(x = "Dimension 1", y = "Dimension 2") 
    p <- p + geom_point(data = mca.dim.ind, aes(x = `Dim 1`, y = `Dim 2`), color = "red", alpha = ind.alpha1) #cloud of individuals
    p <- p + theme_sjplot()
    p1 <- p + theme(
                    panel.grid.major.y = element_line(color = "grey90"),
                    panel.grid.major.x = element_line(color = "grey90"),
                    axis.title = element_text(size = 18), 
                    axis.text = element_text(size = 18),
                    axis.line.x = element_blank(),
                    axis.line.y = element_blank(),
                    legend.position = "none"
                    )
    p <- ggplot(data = mca.dim.vars, mapping = aes(x = `Dim 3`, y = `Dim 4`))
    p <- p + geom_text(aes(label = var.names), size = text.size1, color = "blue", check_overlap = TRUE) #cloud of modalities
    p <- p + geom_hline(yintercept = 0, colour = "gray45") 
    p <- p + geom_vline(xintercept = 0, colour = "gray45") 
    p <- p + labs(x = "Dimension 3", y = "Dimension 4") 
    p <- p + geom_point(data = mca.dim.ind, aes(x = `Dim 1`, y = `Dim 2`), color = "red", alpha = ind.alpha1) #cloud of individuals
    p <- p + xlim(-1.5, 2)
    p <- p + theme_sjplot()
    p2 <- p + theme(
                    panel.grid.major.y = element_line(color = "grey90"),
                    panel.grid.major.x = element_line(color = "grey90"),
                    axis.title = element_text(size = 18), 
                    axis.text = element_text(size = 18),
                    axis.line.x = element_blank(),
                    axis.line.y = element_blank(),
                    legend.position = "none"
                    )
    p.in <- fviz_eig(mca.res, ncp = 8, linecolor = "red")  #inset bar plot showing eigenvalue distribution
    p.in <- p.in + labs(y = "", x = "", title = "") 
    p.in <- p.in + theme_sjplot2()
    p.in <- p.in + theme(
                    plot.background = element_rect(colour = "grey60", size = 0.25),
                    panel.grid.major.y = element_blank(),
                    panel.grid.major.x = element_blank(),
                    )
    p3 <-  p1 / p2 #combining plots using patchwork
    p4 <- p3 + inset_element(p.in, 0.8, 0, 1, 0.5) #inset plot using patchwork
    p4
    save_plot(here("Plots", "Regular", "genre-mca-regular.png"), 
              width = 35, height = 40)
```