### 2023-08-05 blah.tex

        Measures to values: \\
    \begin{itemize}
        \item[--] Education.- Ordinal variable: (1) Less than high school, (2) High school, (3) Associate's Degree, (4) Some college, (5) Bachelor's degree, (6) MA degree, (7) Doctoral or Professional degree.
        \item[--] Parent's College.- Binary variable: At least one parent obtained a college degree (1).
        \item[--] Age.- Ordinal variable: (1) 18--19, (2) 20--25, (3) 25--29, (4) 30--34, (5) 35--39, (6), 40--44, (7), 45--49, (8), 50--54, (9) 55--59, (10), 60--64, (11) 65--69, (12) 70--74, (13) 75--79, (14) 80+.
    \end{itemize}