### 2023-08-05 title-page.tex

\documentclass[a4paper,12pt]{extarticle}
\usepackage[utf8]{inputenc}
\usepackage{blindtext}
\usepackage{geometry}
    \geometry{a4paper}

\title{From Macrogenres to Microgenres via Relationality\footnote{Reserved for acknowledgements.}}

\author{Omar Lizardo \\ \small Department of Sociology \\ \textit{\small University of California, Los Angeles}\thanks{375 Portola Plaza, 265 Haines Hall, Los Angeles California, 90095.} \\ \small olizardo@soc.ucla.edu}

\begin{document}
\maketitle

\end{document}