### 2023-08-05 bio.tex

\documentclass[a4paper,12pt]{extarticle}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
    \geometry{a4paper}
\title{Author's Biography}

\begin{document}
\maketitle
Omar Lizardo (He/Him) is the LeRoy Neiman Term Chair Professor in the Department of Sociology at the University of California, Los Angeles. He studies Culture, Cognition, Networks, Consumption, Institutions, Organization, and Social Theory. His work has appeared in such journals as \textit{American Sociological Review}, \textit{Social Forces}, \textit{Sociological Forum}, \textit{Sociological Theory}, \textit{Theory \& Society}, \textit{Journal for the Theory of Social Behaviour}, \textit{Poetics}, and \textit{Social Networks}. He is the co-editor of the \textit{Handbook of Classical Sociological Theory} published by Springer and a co-author of the book \textit{Measuring Culture} published by Columbia University Press. He serves on the editorial board of \textit{Social Forces}, \textit{Journal for the Theory of Social Behaviour}, \textit{Theory \& Society}, and \textit{Poetics}. He is also a member of the Board of Reviewing Editors for the journal \textit{Science}, an Associate Editor for \textit{Discover Data}, one of the current lead editors of \textit{Sociological Theory}, and a former chair of the American Sociological Association's Section on the Sociology of Culture. His recent publications have focused on the decline of cultural participation in Southern Europe in the wake of the Eurozone crisis, methods for analyzing relational similarities in two-mode networks, and theorizing analytical approaches to the study of cultural phenomena.
\end{document}